---
title: "Working with the R package BCEA"
author: Gianluca Baio
date: 17 June 2021
institute: "[Department of Statistical Science](https://www.ucl.ac.uk/statistics/) | University College London"
params: 
   conference: "ISPOR Short courses"
   location: "Internetville"
   short_title: "HTA and BCEA"
output:
  xaringan::moon_reader:
    includes: 
       in_header: "assets/latex_macros.html" 
       # This line adds a logo based on the format selected in the file 'assets/include_logo.html'
       # NB: the actual options (eg placement of the logo and actual logo file) can be changed there
       # after_body: "assets/insert-logo.html"
    seal: false
    yolo: no
    lib_dir: libs
    nature:
      beforeInit: ["https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: yes
      highlightSpans: true
      countIncrementalSlides: no
      ratio: '16:9'
      titleSlideClass:
      - center
      - middle
    self_contained: false 
    css:
    - "assets/ucl-powerpoint.css"
---

```{r echo=F,message=FALSE,warning=FALSE,comment=NA}
# Sources the R file with all the relevant setup and commands
source("assets/setup.R")

# Stuff from 'xaringanExtra' (https://pkg.garrickadenbuie.com/xaringanExtra)
# This allows the use of panels (from 'xaringanExtra')
xaringanExtra::use_panelset()
# This allows to copy code from the slides directly
xaringanExtra::use_clipboard()
# This freezes the frame for when there's a gif included
xaringanExtra::use_freezeframe()

# Defines the path to the file with the .bib entries (in case there are references)
bibfile=ReadBib("~/Dropbox/Perso/Office/CV/mypubs.bib",check = FALSE)

# Removes '>' from R printouts
knitr::opts_chunk$set(prompt=FALSE)

# Installs the old version (CRAN) of BCEA so it can be used NB: ONLY DO THIS ONCE! 
#tmp_lib="~/Dropbox/HE/Ispor/Workshop_Modelling_R/BCEA/tmp"
#dir.create(tmp_lib)
#install.packages("BCEA",lib=tmp_lib)
tmp_lib="~/Dropbox/HE/Ispor/Workshop_Modelling_R/BCEA_talk/tmp"
```

class: title-slide

# `r rmarkdown::metadata$title`

## `r rmarkdown::metadata$author`

### `r rmarkdown::metadata$institute`    

.title-small[
`r icon::icon_style(icon::fontawesome("envelope",style = "solid"),scale=.8,fill="#00acee")`  [g.baio@ucl.ac.uk](mailto:g.baio@ucl.ac.uk)
`r icon::icon_style(icon::fontawesome("firefox"),scale=.8,fill="#EA7600")`  [http://www.statistica.it/gianluca/](http://www.statistica.it/gianluca/)
`r icon::icon_style(icon::fontawesome("firefox"),scale=.8,fill="#EA7600")`  [https://egon.stats.ucl.ac.uk/research/statistics-health-economics/](https://egon.stats.ucl.ac.uk/research/statistics-health-economics/)
`r icon::icon_style(icon::fontawesome("github"),scale=.8,fill="black")`  [https://github.com/giabaio](https://github.com/giabaio)
`r icon::icon_style(icon::fontawesome("github"),scale=.8,fill="black")`  [https://github.com/StatisticsHealthEconomics](https://github.com/StatisticsHealthEconomics)
`r icon::icon_style(icon::fontawesome("twitter"),scale=.8,fill="#00acee")`  [@gianlubaio](https://twitter.com/gianlubaio)     
]

### `r rmarkdown::metadata$params$conference`, `r rmarkdown::metadata$params$location` 
<!-- Can also separate the various components of the extra argument 'params', eg as in 
### `r paste(rmarkdown::metadata$params, collapse=", ")`
-->

`r date`

<!-- This adds a footer (optional and with other possibilities...) -->
.footer-left[
`r samptux()` <span style="position: relative; bottom: 5px; color: #D5D5D5;"> &nbsp; &copy; Gianluca Baio (UCL)</span>
]

---

layout: true
.footer-left[
`r samptux()` <span style="position: relative; bottom: 5px; color: #D5D5D5;"> &nbsp; &copy; Gianluca Baio (UCL)</span>
]
<!-- Can also include social media icons & hyperlinks -->
.footer-social[
`r add_twitter()` `r add_email()` `r add_website()`
]

<!-- Can also add a center footer, eg to include the title of the talk -->
.footer-center[
`r rmarkdown::metadata$params$short_title`
]
<!-- And a right footer, to include the date -->
.footer-right[
`r rmarkdown::metadata$params$conference`, `r short_date`
]

---

# `BCEA` 

## An `R` package for (Bayesian) cost-effectiveness analysis

### What is `BCEA` not?

- `BCEA` is **not** a package to automatically run a Bayesian analysis
   - It cannot build the health economic model for you
   - It does not prepare the data to be used in the model
   - It does not automatically run the MCMC simulations
   - It does not choose the prior distributions for you

--

### So what *is* it then?

- `BCEA`  provides a set of specific functions to systematically post-process the output of a Bayesian health economic model

- Uses `R` [http://cran.r-project.org/](http://cran.r-project.org/)
   -  Very good at interacting with standard MCMC software
   - `BUGS`: [www.mrc-bsu.cam.ac.uk/bugs/winbugs/contents.shtml](www.mrc-bsu.cam.ac.uk/bugs/winbugs/contents.shtml)
   - `JAGS`: [www.mcmc-jags.sourceforge.net/](www.mcmc-jags.sourceforge.net/)    
   - `rstan`: [https://mc-stan.org/users/interfaces/rstan](https://mc-stan.org/users/interfaces/rstan)
   - .red[**Free**] and there is a very large community of contributors
   - Specifically designed for statistical analysis and has very good graphical capabilities

---

# `BCEA`

## A `R` package for (Bayesian) cost-effectiveness analysis

### `BCEA` and its use directly in `R` are designed with these objectives in mind

1. Checking the model assumptions

   - Do we mean what we mean (eg in terms of PSA simulations)?...
   - Simulation error (especially, **but not only**, for a Bayesian approach)

--

2. Produce the base-case economic evaluation
   - What’s the most cost-effective intervention, given current evidence?
   - Cost-effectiveness plane, Expected Incremental Benefit (as a function of $k$),...

--
 
3. Perform uncertainty analysis
   - Standard PSA (mandatory): Cost-effectiveness Plane, CEAC, ...
   - Fairly easy (but not always used): CEAF
   - More advanced/"too difficult" (rarely used): EVP(P)I/EVSI

--

4. Standardised reporting
   - Graphical tools (use **excellent** `R` facilities)

    
---

count: false

# `BCEA` 

## An `R` package for (Bayesian) cost-effectiveness analysis

```{r bcea_scheme,engine='tikz', echo=F, out.width="85%",opts=list(width="75%",title="INSERT TEXT HERE"),eval=FALSE}
\begin{tikzpicture}

\draw(0,0) node[align=center,rectangle,rounded corners=2ex,draw,fill=blue!40,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](1){\texttt{bcea}};

\draw(-2,-2.9) node[align=center,rectangle,rounded corners=2ex,draw,fill=orange,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.6cm,minimum height=2.6cm,text depth=2.0cm](2){\texttt{plot}};

\draw(-2.0,-3.1) node[align=left,rectangle,rounded corners,draw,fill=red,font=\sffamily\fontsize{6}{7}\selectfont,text width=1.7cm](18){
\tiny$\bullet$ \texttt{ceplane.plot} \\
\tiny$\bullet$ \texttt{contour} \\
\tiny$\bullet$ \texttt{contour2} \\
\tiny$\bullet$ \texttt{ceac.plot} \\
\tiny$\bullet$ \texttt{evi.plot} \\
\tiny$\bullet$ \texttt{eib.plot} \\
\tiny$\bullet$ \texttt{ib.plot} \\
};

\draw(1.5,-3) node[align=center,rectangle,rounded corners=2ex,draw,fill=orange,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](3){\texttt{summary}};

\draw(3,2.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=blue!40,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](4){\texttt{evppi}};

\draw(0,2.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=red,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](5){\texttt{CreateInputs}};

\draw(-3,2.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=red,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](6){\texttt{struct.psa}};

\draw(5,3.0) node[align=center,rectangle,rounded corners=2ex,draw,fill=orange,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](7){\texttt{plot}};

\draw(5,2.0) node[align=center,rectangle,rounded corners=2ex,draw,fill=red,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](8){\texttt{diag.evppi}};

\draw(-3,0.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=red,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](9){\texttt{multi.ce}};

\draw(-5,1) node[align=center,rectangle,rounded corners=2ex,draw,fill=red,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](10){\texttt{mce.plot}};

\draw(-5,0) node[align=center,rectangle,rounded corners=2ex,draw,fill=red,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](11){\texttt{ceaf.plot}};

\draw(-5,-1.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=red,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](12){\texttt{ceef.plot}};

\draw(3,.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=blue!40,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](13){\texttt{CEriskav}};

\draw(5,.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=orange,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](14){\texttt{plot}};

\draw(3,-1.5) node[align=center,rectangle,rounded corners=2ex,draw,fill=blue!40,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](15){\texttt{mixedAn}};

\draw(5,-1) node[align=center,rectangle,rounded corners=2ex,draw,fill=orange,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](16){\texttt{plot}};

\draw(5,-2) node[align=center,rectangle,rounded corners=2ex,draw,fill=orange,font=\sffamily\fontsize{7}{8}\selectfont,minimum width=1.2cm,minimum height=.6cm](17){\texttt{summary}};

\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (2.70) -- (1.210);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (1.240) -- (18.0);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (3.90) -- (1.300);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (1.160) -- (9.360);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (1.140) -- (6.310);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (1.50) -- (4.220);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (1.20) -- (13.west);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (1.340) -- (15.west);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (1.west) -- (12.east);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (7.west) -- (4.15);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (4.340) -- (8.west);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (16.west) -- (15.15);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (17.west) -- (15.345);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (14.west) -- (13.east);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (5.east) -- (4.west);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (5.west) -- (6.east);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (9.160) -- (10.east);
\draw [->,>=latex,shorten >=0pt,auto,node distance=1cm,ultra thin] (9.200) -- (11.east);

\end{tikzpicture}
```

`r include_fig("bcea_scheme-1.png",width="65%",title="")`

`r vspace("-50px")`
 
.small[
`r icon::fontawesome("firefox")` [www.statistica.it/gianluca/BCEA](www.statistica.it/gianluca/BCEA)     
`r icon::icon_style(icon::fontawesome("github"),scale=.8,fill="black")`  [https://github.com/giabaio/BCEA](https://github.com/giabaio/BCEA)
]

---

class: empty-slide

.huge[Part I]
# Using `BCEA` to summarise outputs of an economic model

---

# How does `BCEA` work?

.panelset[
.panel[
.panel-name[Installation]
- `BCEA` is available from [CRAN](https://cran.r-project.org/)
   - Current *stable* version: 2.3-1.1 (26 August 2019)

- But it is also under constant development in the [GitHub repository](https://github.com/giabaio/BCEA)
   - Current *stable* version: 2.3-2 (30 January 2020)
   - Current *development* version: 2.4-1 (5 February 2021) &ndash; **major** update (soon to be released!)

```{r howwork1,eval=FALSE}
# Install BCEA (only required once and needs an internet connection!). 

# You can either get the "official" version from CRAN
install.packages("BCEA")

# Or the development version from the GitHub repository
devtools::install_github("giabaio/BCEA")               # stable version (2.3-2)

devtools::install_github("giabaio/BCEA",ref="devel")   # development version (2.4-1)
```

`r vspace("30px")`
NB: The beauty of the GitHub version is that it can be updated on the fly and be immediately available for users!
]

.panel[
.panel-name[Using BCEA]
```{r howwork2,eval=TRUE,echo=FALSE}
library(BCEA,lib=tmp_lib)
data(Vaccine)
inp = CreateInputs(vaccine, print.lincom=FALSE)
```
```{r howwork3,eval=FALSE}
library(dplyr)  # (Not necessary - helpful for data manipulation!)

library(BCEA)   # Then loads the package (so you can access its functions)
data(Vaccine)   # Loads an example dataset 
```

`r vspace("40px")`
- The "Vaccine" example is a fictional cost-effectiveness model for and influenza vaccine, based on evidence synthesis (and a real case)

- 2 treatment options ("Standard of care" vs "Vaccination") and overall 63 parameters

- Discussed in details in `r icon::fontawesome("book-open")` [Baio et al, 2017](http://www.statistica.it/gianluca/book/bcea/) and `r icon::academicons$pubmed` [Baio and Dawid, 2011](https://pubmed.ncbi.nlm.nih.gov/21930515/)

- In this case, PSA simulations obtained from a full Bayesian model, but could be done in a spreadsheet and imported into `R`
]

.panel[
.panel-name[Show. Me. The. Data]
`r vspace("-10px")`
```{r howwork4,eval=FALSE}
# Creates a matrix with the underlying model simulations
inp = CreateInputs(vaccine, `print.lincom=FALSE`)           `# NB: this option will be slightly different in 2.4-1!`
# Visualise the output
inp$mat `%>% as_tibble()`                                   `# "piping" ('%>%') and formatting nicely`
```

```{r howwork5,echo=FALSE}
inp$mat %>% as_tibble()
```
]

.panel[
.panel-name[Economic model]

```{r howwork6,eval=FALSE}
# Defines the number of simulations considered
n.sims=`inp$mat %>% nrow()`                             `# applies the function 'nrow' (number of rows) to the object 'inp$mat'`

                                                       `# NB: in R 4.1.0, can also use 'native' pipe ('|>')`
                                                       `# (probably a bit quicker, but in most cases, may be immaterial...)`

# Aggregates the model inputs to compute (e,c)
QALYs.inf = QALYs.pne <- QALYs.hosp <- QALYs.adv <- QALYs.death <- matrix(0,n.sims,2)
for (t in 1:2) {
  QALYs.inf[,t] = ((Infected[,t,1] + Infected[,t,2])*omega[,1]/365)/N
  QALYs.pne[,t] = ((Pneumonia[,t,1] + Pneumonia[,t,2])*omega[,4]/365)/N
  QALYs.hosp[,t] = ((Hospital[,t,1] + Hospital[,t,2])*omega[,5]/365)/N
  QALYs.death[,t] = ((Death[,t,1] + Death[,t,2])*omega[,6])/N
}
QALYs.adv[,2] = (Adverse.events*omega[,7]/365)/N
e = -(QALYs.inf + QALYs.pne + QALYs.adv + QALYs.hosp + QALYs.death) + ...
```

`r vspace("40px")`
- NB: The data stored in the `Vaccine` object (built-in in `BCEA`) already contains the objects `e,c` that can be used to run the decision analysis...

- So, this step is actually not needed (but documented in `r icon::fontawesome("book-open")` [Baio et al, 2017](http://www.statistica.it/gianluca/book/bcea/))
]

.panel[
.panel-name[Cost & effects]
```{r howwork7,eval=FALSE}
cbind(e,c) %>% as_tibble(`.name_repair="universal"`)     `# ensures that the columns are named` 
```
```{r howwork8,echo=FALSE}
cbind(e,c) %>% as_tibble(.name_repair="universal")
```

`r vspace("40px")`

- These calculations *can* be done also in a spreadsheet (nothing more than algebra, **once you have the simulations**)
]
]

---

count: false
# How does `BCEA` work?

- At this point, we are ready to call the function `bcea` that runs the economic analysis, for example something like 

```{r howwork9}
treats = c("Status quo","Vaccination")
m = bcea(e=e,c=c,ref=2,interventions=treats,Kmax=50000)
```

- The inputs to the function are
   - `e`: a **matrix** containing the simulations for the clinical benefits (that is $n_{\rm{sim}}\times n_{\rm{int}}$ values)
   - `c`: a **matrix** containing the simulations for the costs (that is $n_{\rm{sim}}\times n_{\rm{int}}$ values)
   -  `ref`: an indication of which intervention is to be taken as reference (default: the intervention in the first column of `e` or `c`)
   - `interventions`: a vector of labels for the interventions being compared 
   - `Kmax`: the maximum value of $k$, the parameter of willingness to pay


--


- The output is an object `m` containing several elements
```{r howwork10}
names(m)
```

---

count: false
# How does `BCEA` work?

Can visualise the output in various formats (tables/graphs)

```{r howwork11}
# The 'summary' "method" produces a tabular output
summary(m)
```

---

count: false
# How does `BCEA` work?

Can visualise the output in various formats (tables/graphs)

```{r howwork12,opts=list(title="",width="40%")}
# The 'plot' "method" produces a *specific* version of graphical output
plot(m)
```

---

count: false
# How does `BCEA` work?

```{r howwork13,opts=list(title="",width="45%")}
ceplane.plot(m,wtp=20000,xlim=c(-.002,.002),ylim=c(-10,20))
```

---

count: false
# How does `BCEA` work?

```{r howwork14a,eval=FALSE}
# Using 'ggplot', you can go crazy with customisation...
ceplane.plot(m,wtp=10000,`graph="gg"`,point_colors="blue",point_sizes=.8,area_color="springgreen3")

```

```{r howwork14,opts=list(title="",width="45%"),echo=FALSE}
# Using 'ggplot', you can go crazy with customisation...
ceplane.plot(m,wtp=10000,graph="gg",point_colors="blue",point_sizes=.8,area_color="springgreen3")
```

`r vspace("-40px")`
.small[`r icon::fontawesome("firefox")` [https://ggplot2.tidyverse.org/](https://ggplot2.tidyverse.org/)]

---

count: false
# How does `BCEA` work?

```{r howwork15,opts=list(title="",width="45%")}
# Plots the Cost-Effectiveness Acceptability Curve
ceac.plot(m)
```

---

count: false
# How does `BCEA` work?

```{r howwork16,opts=list(title="",width="45%")}
# Plots the Expected Value of Partial Information (EVPI)
evi.plot(m)
```

---

# Specialised plots

- Can generate a *contour* plot of the cost-effectiveness plane and estimate the proportion of points in each quadrant

.left-column[
```{r howwork17,opts=list(title="",width="45%"),eval=FALSE}
# "Basic" contourplot
contour(m)
```
]

.right-column[
```{r howwork18,opts=list(title="",width="65%"),echo=FALSE}
# "Basic" contourplot
contour(m)
```
]

---

count: false
# Specialised plots

- The specialised function `contour2` also shows the **sustainability area**

.pull-left[
```{r howwork19,opts=list(title="",width="80%")}
contour2(m)
```
]

.pull-right[
`r vspace("-0px")`
```{r howwork20,opts=list(title="",width="80%")}
contour2(m,wtp=100)
```
]
---

count: false
# Specialised plots
### Cost-effectiveness efficiency frontier

.pull-left[
```{r howwork20bis,opts=list(title="",width="80%")}
ceef.plot(m,print.plot=FALSE)
```
]

.pull-right[
`r vspace("-0px")`
```{r howwork20tris,opts=list(title="",width="80%")}
ceef.plot(m,print.summary=FALSE)
```
]---

# Exporting graphical output

- `R` has excellent graphical facilities and the graphs produced by `BCEA` can be easily exported to many different formats

```{r howwork21,eval=FALSE}
# "Opens" the graphical device
pdf("NAME_OF_THE_FILE",width=`8`,height=`8`)      `# for 'pdf', units are in inches`
# Makes the plot
ceplane.plot(`BCEA_OBJECT`)                     `# of course, specify whatever name you've chosen when creating the object...`
# "Closes" the graphical device
dev.off()


# "Open" the graphical device"
jpeg("NAME_OF_FILE.jpg",width=`480`,height=`480`)  `# for 'jpeg' units are in px`
# Makes the plot
ceplane.plot(BCEA_OBJECT)
# "Closes" the graphical device
dev.off()
```


`r vspace("30px")`
**NB**: `Rstudio` and [`rmarkdown`](https://rmarkdown.rstudio.com/) can do even more &ndash; that's for another time...
---

class: empty-slide

.huge[Part II]
# Advanced use of `BCEA`

---

# Multiple treatment comparisons
## Probabilistic "depression model"

```{r depression1,echo=FALSE}
n.samples<-1000 

# Number and names of treatments
n.treat<-3
t.names<-c("No treatment","CBT","Antidepressant")

# Over a 30 year time horizon
# Costs for recovery, relapse, and no recovery
c.rec<-rnorm(n=n.samples, mean=1000, sd=50)
c.rel<-rnorm(n=n.samples, mean=2000, sd=100)
c.norec<-rnorm(n=n.samples, mean=2500, sd=125)
# Over a 30 year time horizon
# QALYs for recovery, relapse, and no recovery
q.rec<-rnorm(n=n.samples, mean=26, sd=2)
q.rel<-rnorm(n=n.samples, mean=23, sd=3)
q.norec<-rnorm(n=n.samples, mean=20, sd=4)

# Probabilities of recovery and relapse following recover
p.rec<-p.rel<-matrix(nrow=n.samples, ncol=n.treat)

# Probabilities for no treatment
p.rec[,1]<-rbeta(n=n.samples, shape1=6, shape2=200)
p.rel[,1]<-rbeta(n=n.samples, shape1=2, shape2=100)

# Probabilities for CBT
# Probability of recovery higher than on no treatment
p.rec[,2]<-rbeta(n=n.samples, shape1=6, shape2=130)
# Probability of relapse lower than on no treatment
p.rel[,2]<-rbeta(n=n.samples, shape1=2, shape2=200)

# Probabilities for antidepressant
# Probability of recovery lower than no treatment or CBT
p.rec[,3]<-rbeta(n=n.samples, shape1=6, shape2=120)
# Probability relase lower than no treatment, higher than CBT
p.rel[,3]<-rbeta(n=n.samples, shape1=2, shape2=120)
# Cost of treatment 
c.treat<-t(matrix(rep(c(0,300,30),n.samples),ncol=n.samples,nrow=3))
# Total cost and effects 
costs<-c.treat+p.rec*(1-p.rel)*c.rec+p.rec*p.rel*c.rel+(1-p.rec)*c.norec
effects<-p.rec*(1-p.rel)*q.rec+p.rec*p.rel*q.rel+(1-p.rec)*q.norec
```

- Fictional model comparing antidepressants to cognitive behaviour therapy (CBT) and no treatment in people with depression

- Statistical modelling based on evidence synthesis
   - Benefits: based on QALYs
   - Costs: associated with treatments and various resources use
   
- Economic modelling: two matrices with relevant population summaries
   - `effects`
   - `costs`
   
`r vspace("30px")`
- NB: The details of the actual modelling are *not* important for the purposes of demonstrating the example...
   
---

count: false
# Multiple treatment comparisons
## Probabilistic "depression model"

.pull-left[
```{r depression2,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="60%"),eval=FALSE}
# Intervention labels
t.names<-c("No treatment","CBT","Antidepressant")

# "Standard" analysis: pairwise comparisons
depression.bcea = bcea(effects,costs,
                       interventions=t.names,`ref=3`)  
# `the third intervention is the reference`

# Plots the results
plot(depression.bcea)
```
]
.pull-right[
```{r depression3,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="90%"),echo=FALSE}
# Intervention labels
t.names<-c("No treatment","CBT","Antidepressant")

# "Standard" analysis: pairwise comparisons
depression.bcea = bcea(effects,costs,interventions=t.names,ref=3)

# Plots the results
plot(depression.bcea)
```

]

---

count: false
# Multiple treatment comparisons
## Probabilistic "depression model"

.pull-left[
```{r depression4,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="60%"),eval=FALSE}
# For `multiple treatment comparison`
depression.multi.ce = `multi.ce`(depression.bcea)

# Specialised plot method
`mce.plot`(depression.multi.ce,pos=c(1,0.8),
           graph=c("base","ggplot2"))
```
]
.pull-right[
```{r depression5,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="90%"),echo=FALSE}
depression.multi.ce = multi.ce(depression.bcea)
mce.plot(depression.multi.ce,pos=c(1,0.8),graph=c("base","ggplot2"))
```

]

---

count: false
# Multiple treatment comparisons
## Probabilistic "depression model"

.pull-left[
`r vspace("20px")`
- Can use `ggplot` to customise the graph 
```{r depression6,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="60%"),eval=FALSE}
mce.plot(depression.multi.ce,pos=c(1,1),graph="ggplot2") + 
  ggplot2::stat_summary(fun=max, geom="line", 
                        colour="grey25", alpha=.3, lwd=2.5)
```
]
.pull-right[
```{r depression7,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="90%"),echo=FALSE}
mce.plot(depression.multi.ce,pos=c(1,1),graph="ggplot2") + 
  ggplot2::stat_summary(fun=max, geom="line", colour="grey25", alpha=.3, lwd=2.5)
```
]

---

count: false
# Multiple treatment comparisons
## Probabilistic "depression model"

.pull-left[
`r vspace("20px")`
- Can use a specialised plotting method for the *cost-effectiveness acceptability **frontier***
```{r depression8,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="60%"),eval=FALSE}
# Specialised plot
ceaf.plot(depression.multi.ce)
```
]
.pull-right[
```{r depression9,fig.width=8,fig.height=8,fig.align='center',message=FALSE,warning=FALSE,opts=list(width="90%"),echo=FALSE}
ceaf.plot(depression.multi.ce)
```
]

---

# `BCEAweb`

-  Inspired by similar projects &ndash; eg [SAVI](http://savi.shef.ac.uk/SAVI/)

- Create a web interface to use BCEA without even opening R (or even having it installed on your computer!)
`r vspace("40px")`

--

- Typical work flow
   1. Design the economic model (eg Markov model, decision tree, ...)
   2. Run the statistical analysis to estimate the quantities of interest (eg survival analysis, evidence synthesis, ...)
   3. Run the economic model and obtain "PSA samples"     

--
`r vspace("40px")`
  
   4. Upload "PSA samples", including values for $(e,c)$ to `BCEAweb`
   5. Use `BCEA` in the background to do **all** the economic analysis
   6. Create reports that can be used as the basis for papers, reimbursement files, ...


---

count: false
# `BCEAweb`

```{r eval=FALSE}
# Creates a matrix with the underlying model simulations
inp = CreateInputs(vaccine, print.lincom=FALSE)

# Runs BCEAweb
BCEAweb(e=e,                 # matrix of simulations for the effectiveness
        c=c,                 # matrix of simulations for the costs
        parameters=inp$mat   # matrix of simulations for all the model parameters
       )
```

`r vspace("30px")`

- `BCEAweb` exists as a standalone webapp
   
   - Can access it `r icon::fontawesome("firefox")`  [here](https://egon.stats.ucl.ac.uk/projects/BCEAweb/) 
   
- **Or**, you can launch your own "local" version from the `BCEA` package (as in the code above)!
   
   - This will launch a web page from which you can manipulate your output (`r icon::fontawesome("arrow-circle-right")` Live Demo)

---
class: thankyou-michelle 
